My title

### 2024-02-19 paper/paper.qmd
committing changes to paper.qmd Feb 18 @ 1:08.
Committing all changes made to paper.qmd and references.bib up unitl this point (Feb 18 @ 8:45pm).

The intersection of race and criminal justice policy in the United States has long been a point of contention, marked by pronounced divergences in perception between Black and White Americans. Prevailing assumptions suggest a steadfastly punitive stance in U.S. public opinion towards criminal justice issues, with such punitiveness among Whites often attributed to underlying anti-Black prejudice @BoboJohnson2004. This phenomenon is particularly pronounced in the context of the war on drugs, where drug epidemics are often portrayed negatively within society, with a tendency to associate such crises predominantly with non-White and minority communities. This portrayal not only reinforces stigmatizing stereotypes but also influences public opinion and policy.

@@

Raychaudhuri, Mendelberg, and McDonough's study investigates the impact of media coverage on public support for drug treatment policies, focusing on how the racial framing and valence (positive or negative portrayal) of opioid abuse influence White Americans' policy views. Unlike previous drug epidemics portrayed negatively and associated with non-White users, the opioid crisis is often framed sympathetically and as predominantly affecting White Americans. This sympathetic portrayal is thought to contribute to the broad public support for treating opioid abuse as a public health issue rather than a criminal one @Raychaudhuri2022.

The goal of the study was to asses whether the sympathetic portrayal of opioid users, predominantly White, influences public support for health-centered approaches over punitive measures and if such sympathy extends to Black individuals or remains racially selective. The study discovered that when opioid users, both White and Black, were shown in a sympathetic light, White people's support for treatment options rose. However, this supportive attitude was noticeably stronger when the users depicted were White, pointing to a bias in empathy based on race. Essentially, White Americans showed a preference for offering treatment to individuals of their own racial background. On the other hand, portrayals lacking sympathy didn't really change opinions, regardless of the user's race @Raychaudhuri2022.

@@

# Features

@@

@fig-interactions visualizes the interaction effects between political ideology (conservative or liberal) and media framing (sympathetic or unsympathetic, towards white or black opioid users) on the support for policy regarding opioid abuse treatment.

@@

tidy_model <- broom::tidy(model_ideology_frame) %>%
  filter(term %in% c("liberal:unsympathetic_white", 
                     "liberal:unsympathetic_black", 
                     "liberal:sympathetic_white", 
                     "liberal:sympathetic_black", 
                     "conservative:unsympathetic_white", 
                     "conservative:unsympathetic_black", 
                     "conservative:sympathetic_white", 
                     "conservative:sympathetic_black"))

@@

>>>>

The intersection of race and criminal justice policy in the United States has long been a point of contention, marked by notable divergences in perception between Black and White Americans. Predominant assumptions suggest a steady punitive stance in U.S. public opinion towards criminal justice issues, with such punitiveness among Whites often attributed to underlying anti-Black prejudice @BoboJohnson2004. This phenomenon is particularly pronounced in the context of the war on drugs, where drug epidemics are often portrayed negatively within society, with a tendency to associate such crises predominantly with non-White and minority communities. This portrayal not only reinforces stigmatizing stereotypes but also influences public opinion and policy.

@@

Raychaudhuri, Mendelberg, and McDonough's study investigates the impact of media coverage on public support for drug treatment policies, focusing on how the racial framing and valence of opioid abuse influence White Americans' policy views. Unlike previous drug epidemics portrayed negatively and associated with non-White users, the opioid crisis is often framed sympathetically and as predominantly affecting White Americans. This sympathetic portrayal is thought to contribute to the broad public support for treating opioid abuse as a public health issue rather than a criminal one @Raychaudhuri2022.

The study revealed that when opioid users, both White and Black, were shown in a sympathetic light, White people's support for treatment options rose. However, this supportive attitude was noticeably stronger when the users depicted were White, pointing to a bias in empathy based on race. Essentially, White Americans showed a preference for offering treatment to individuals of their own racial background. On the other hand, portrayals lacking sympathy didn't really change opinions, regardless of the user's race @Raychaudhuri2022.

@@

@@

@fig-interactions replicates the forests plots used in Raychaudhuri, Mendelberg, and McDonough's study. This replication uses the new model outlined in this paper by visualizing the interaction effects between political ideology (conservative or liberal) and media framing (sympathetic or unsympathetic, towards white or black opioid users) on the support for policy regarding opioid abuse treatment.

@@

tidy_model <- broom::tidy(model_ideology_frame) |>
  filter(term %in% c("liberal:unsympathetic_white", 
                     "liberal:unsympathetic_black", 
                     "liberal:sympathetic_white", 
                     "liberal:sympathetic_black", 
                     "conservative:unsympathetic_white", 
                     "conservative:unsympathetic_black", 
                     "conservative:sympathetic_white", 
                     "conservative:sympathetic_black"))

@@

# Discussion
The interest to examine the interaction effects between political ideology (conservative or liberal) and media framing (sympathetic or unsympathetic, towards white or black opioid users) on the support for policy regarding opioid abuse treatment was heavily influenced by the reading of Pyra M, Taylor B, Flanagan E, Hotton A, Johnson O, Lamuda P, Schneider J, and Pollack H's study titled "Support for evidence-informed opioid policies and interventions: The role of racial attitudes, political affiliation, and opioid stigma" @Pyra2022. The significance of the study was that it emphasized that despite the opioid epidemic being portrayed as less racially and politically divisive compared to previous drug crises, underlying racial and political attitudes remain influential @Pyra2022. Their study revealed that Democratic Party affiliation (political ideological outlook of "liberal" in the sense of this paper) correlated with greater support for these measures and mandatory treatment, possibly as an alternative to more punitive approaches.

Building upon the insights from Pyra et al. and Raychaudhuri, Mendelberg, and McDonough, this discussion contributes to the understanding of how political ideologies intertwine with media framings that influence public support for opioid abuse treatment policies.

The null hypothesis ($H_0$) stated that there would be no interaction effect between political ideology and media framing on support for opioid abuse treatment policy. However, @fig-interactions shows varying estimates across different combinations of political ideology and media framing. This suggests that there may be an interaction effect. This is because the estimates for both liberal and conservative ideologies when paired with certain sympathetic framings are positive and do not overlap with the zero line, indicating a significant effect. This suggests that there is indeed some interaction effect, which leads to the rejection of the null hypothesis.

$H_1$ stated there would be a significant interaction effect between political ideology and sympathetic media framing, with liberals showing a stronger positive response towards treatment-focused policies than conservatives. The graph indicates that both liberals and conservative have a positive estimate for both `:sympathetic_white` and `:sympathetic_black`. This suggests that liberals and conservatives indeed show more support for treatment-focused policies when exposed to sympathetic narratives, regardless of the race of the opioid user. However, the liberal estimate for `:sympathetic_white` is the only variable to not overlap with zero, a sign of a strong positive response towards treatment-focused policies when framed in a white, sympathetic lens. This aids in supporting $H_1$, however, $H_1$ stated that liberals would show a stronger positive response towards treatment-focused policies than conservatives when exposed to sympathetic narratives, regardless of the users race, and this is only true when discussing sympathetic narratives toward white opiod users.

The confidence intervals of the estimates for `conservative:sympathetic_white`, `conservative:sympathetic_black`, and `liberal:sympathetic_black` all cross the zero point, meaning that while their point estimates are positive, suggesting a positive effect, there is not enough statistical evidence to conclude that the effect is significantly different from zero at the chosen level of confidence (95%).

$H_2$ stated there would be a significant interaction effect between political ideology and unsympathetic media framing, expecting that liberals would be less swayed by unsympathetic narratives towards punitive measures. However, `conservative:unsympathetic_black` is the only unsympathetic media frame to show a positive effect and whose confidence intervals do not cross the zero point, disproving this hypotheses. `liberal:unsympathetic_white, liberal:unsympathetic_black, and 'conservative:unsympathetic_white' all show a positive effect, however, their confidence intervals all cross the zero point. at this point it is acceptable to reject $H_2$, as the data does not show a clear differential impact of unsympathetic framing between the ideologies.

### Accounting for Bias
Accounting for bias is a critical component of research arise naturally when collecting quantitative and qualitative data. Raychaudhuri, Mendelberg, and McDonough took important steps to mitigate bias in their experiments by pre-registering their methods and employing a fully controlled design with realistic news stories. This approach helps prevent confirmation bias, meaning that the authors were not influenced by their own beliefs and values to unintentionally back their hypotheses because of the way their studies were conducted. While the authors did present measures to control for bias, there are still potential areas where bias could be present. This could include sampling bias as the authors make no mention to whether the study captures a diverse cross-section of White Americans in terms of factors like socioeconomic status, education level, or demographics.

Other accounts for bias could be present in the media content used for the study. The authors state that the generated news stories used in the study were constructed based on realistic, pre-existing new stories. Therefore, If the realistic news stories used in the study are based on actual content that already has a bias, then the study could inadvertently measure the effects of existing media bias rather than pure racial or valence framing effects. Furthermore, the wording of news stories used in the experiment could influence respondents' answers. Subtle language cues can lead to different interpretations, affecting the study's outcomes. In fact, the entire piece of media is left to interpretation.

### Limitations
While the research offers significant insights into the impact of media framing on policy views, it focuses primarily on White Americans' responses. Future research could explore how different racial groups respond to such framing and whether the effects observed in this study hold across diverse demographics. Exploring the effects across multiple racial groups could aid in strengthening the papers conclusions. Therefore, being a reproduction, this paper faces the same limitation. The potential existence of sampling bias discussed above can also be seen as a limitation and therefore effects the integrity of this paper.

# Future Research 
Future research could look into the psychological mechanisms that may explain why and how political ideologies modulate responses to media framing of opioid users. It is also imperative to understand the role of individual identity, such as race and personal experience with opioid abuse, in shaping one's reaction to policy framing. Could it be that personal connections to the opioid crisis, either through community impact or personal acquaintance with sufferers, are powerful enough to override political predispositions? Personal experience could play a significant role in an individuals views on media framing and policy and seems like a strong future area of research reagrding both this paper and the original.

# Conclusion 
This study contributes to the growing body of literature that seeks to understand the opioid crisis not just as a public health issue, but as a societal issue that is deeply intertwined with race, politics, and media. By replicating and extending the work of Raychaudhuri, Mendelberg, and McDonough, this paper sheds light on the nuanced ways in which these factors interact to shape policy preferences, offering insights that could inform more effective and equitable approaches to addressing the opioid epidemic.
